---
title: "Making the downcaleLST.MDV package"
author: "Maite Lezama Valdes"
date: "11.07.21"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, setup_wd, include=FALSE}
knitr::opts_knit$set(root.dir = '../')
```

```{r development, include=FALSE, warning=FALSE, message=FALSE}
library(testthat)
```


Package was generated with the help of package fusen by Sébastien Rochette. 

```{r description, eval=FALSE}
# Describe your package
fusen::fill_description(
  pkg = here::here(),
  fields = list(
    Title = "Downscaling Land Surface Temperature for the Antarctic Dry Valleys",
    Description = "This package enables users to downscale MOD/MYD11_L2 scenes for daytime hours within the McMurdo Dry Valleys, Antarctica to a 30m spatial resolution",
    `Authors@R` = c(
      person("Maite", "Lezama Valdes", email = "maite.lezama@uni-muenster.de", role = c("aut", "cre"))
    )
  )
)
# Define License with use_*_license()
usethis::use_mit_license("Maite Lezama Valdes")
```

# Read data

The downscaling model contained in this package is called "final_model_downscaling_LST_MDV.RData". 

Auxiliary raster data are in aux.tif and an exemplary MODIS scene to downscale is in MOD11_L2_example.tif, it's from XXX (ID=XXX).

There is also a template file (template.tif) for the 30m resolution and slope.tif and aspect.tif (in degrees) to calculate the incidence angle.

<!-- Store your dataset in a directory named "inst/" at the root of your project -->
<!-- Use it for your tests in this Rmd thanks to `load_all()` to make it available
and `system.file()` to read it in your examples 
-->

```{r development-2}
# Run all in the console directly
# Create "inst/" directory
dir.create(here::here("inst"))
# Example dataset
# file.copy(system.file("nyc_squirrels_sample.csv", package = "fusen"), here::here("inst"))
# Make your dataset file available to the current Rmd
pkgload::load_all(path = here::here(), export_all = FALSE)
# You will be able to read your example data file in each of your function examples and tests as follows

# nyc_squirrels <- read.csv(datafile)
```


# make_ia: Calculate the incidence angle for the MODIS scene

```{r function-make-ia}
#' Calculate incidence angle for MODIS scene to be downscaled
#'
#' @param M_date MODIS scene capturing time in format POSIXct, UTC, just as in the MODIS filename
#' @param outdir directory the incidence angle will be written to
#' @param tmpdir directory for writing temporary raster files
#' @param sl slope in degrees, full MDV file loadable via slope <- raster(system.file("30m_slopeMDV.tif", package = "downscaleLST.MDV"))
#' @param as aspect in degrees, full MDV file loadable via aspect <- raster(system.file("30m_aspectMDV.tif", package = "downscaleLST.MDV")) 
#' @param template template file in format raster
#' @return
#' Incidence Angle raster for the time the MODIS scene to be downscaled was captured
#' @export
#'
#' @examples



make_ia <- function(M_date, outdir, tmpdir = tempdir(),
                    sl=slope, as=aspect, tmplt=template,
                    writeRastertoOutdir=TRUE){
  
  raster::rasterOptions(tmpdir=tmpdir)
  
  print(paste0("starting with ia generation for ",M_date))
  
  doy <- lubridate::yday(M_date)
  h <- lubridate::hour(M_date)
  min <- lubridate::minute(M_date)
  
  doymin <- as.numeric(doy)+((h+min/60)/24)
  
  dateaschar <- gsub(gsub(substring(as.character(M_date), 1,16), pattern="\\s", replacement="_"), 
       pattern=":", replacement="_")
  
  # check whether daylight saving applies
  MODDateNZ <- lubridate::with_tz(M_date, tzone = "Pacific/Auckland")
  if(lubridate::dst(MODDateNZ)==TRUE){
    dstnz <- 60
  } else {
    dstnz <- 0
  }
  
  #  mean latitude and longitude for whole area of interest, i.e. lat=-77.45706,lon= 161.7673
  lat <- (-77.45706)
  lon <- 161.7673
  ia <- solrad::Incidence(DOY = doymin, Lat=lat, Lon=lon, 
                          SLon=180, DS=dstnz, 
                          Slope = sl, Aspect = as)
  
  iares <- raster::resample(ia, tmplt)
  names(iares) <- "ia"
  rm(ia)
  
  if(writeRastertoOutdir==TRUE){
    
      raster::writeRaster(iares, paste0(outdir, "/ia_res", dateaschar, ".tif"), overwrite=T)
      print(paste0("wrote ia", dateaschar,"to ", paste0(outdir, "/ia_res_", dateaschar, ".tif")))
  }
  
  gc()

  return(iares)
  
}

```

The example is based on a small extent, for which template, slope and aspect are provided in the package as "ex_temp.grd", "ex_slope.grd" and "ex_aspect.grd". They are prepared as follows: 


For the full area of interest grid, call: 

full <- raster(system.file("full_size_grids_all_layers.grd", package = "downscaleLST.MDV"))
ex <- raster(system.file("ex_2_all_layers.grd", package = "downscaleLST.MDV"))


```{r examples}

exdate <- as.POSIXct("2018-11-12 13:50:00", tz="UTC")
full <- raster::stack("inst/full_size_grids_all_layers.grd")
ex2 <- raster::stack("inst/ex_2_all_layers.grd")

# for the example extent
ia <- make_ia(M_date = exdate, outdir = here::here("ex_day"), 
              sl=ex2$slope, as=ex2$aspect, tmplt=ex2$template)
# # for the full scene
# ia <- make_ia(M_date = exdate, outdir = here::here("ex_day"),
#               sl=full$slope, as=full$aspect, tmplt=full$template)


```


# downscaleFun: Downscale MODIS scene 

```{r function-downscale}

#' Downscale MODIS scene
#'
#' @param MOD MODIS scene in format raster, will be resampled to template resolution in case it is not already provided as such
#' @param aux auxiliary data in format raster stack as provided by package as aux.grd
#' @param ia incidence angle in format raster as generated by make_ia() 
#' @param template template in format raster as provided by package as template.grd
#' @param modname character, containing the first part of the MODIS scene name at least to derive whether the scene stems from Aqua or Terra satellite 
#' @param outformat default is ".grd" but could also be ".tif" 
#' @param add character, for an additional suffix to the outfile name in order to mark e.g. smaller extens
#' @return
#' downscaled MODIS LST at 30m resolution 
#' @export
#'
#' @examples

downscaleFun <- function(MOD, ia, modname, aux=aux, template=template,
                         writeRastertoOutdir=TRUE, outdir, add=NULL, outformat=".grd"){
  
      if(any(raster::res(MOD) != 30)){
        MOD <- resample(MOD, template, method="bilinear")
      }
      
      TeAqNum <- template
  
      if(grepl("MYD",modname)){
            TeAqNum[] <- 1
            # TeAqNum[1] <- 2
            # TA <- 1
            # corr <- 1
          } else {
            TeAqNum[] <- 2
            # TeAqNum[1] <- 1
            # TA = 2
            # corr <- 2
      
          }
      names(TeAqNum) <- "TeAqNum"
          
      
      # start putting a predictor stack together
      takeaux <- which(names(aux) %in% model_final$finalModel$xNames )
      aux <- aux[[takeaux]]
      names(aux)
      
      predstack <- raster::stack(MOD, aux, ia, TeAqNum)
      
      # prediction is performed on data.frame due to formatting in the model training data
      predstackdf <- data.frame(predstack[])
      predstackdf$soilraster <- factor(predstackdf$soilraster)
      predstackdf$TeAqNum <- factor(predstackdf$TeAqNum)
      predstackdf$landcoverres <- factor(predstackdf$landcoverres)
      names(predstackdf) <- c("Modis", names(predstackdf[2:length(names(predstackdf))]))
      predstackdf <- predstackdf[,model_final$finalModel$xNames] # order columns
      
      # bring new data into trainingData form
      trained_with <- model_final$trainingData[,1:(ncol(model_final$trainingData)-1)]
      if(any(!levels(predstackdf$soilraster)  %in% levels(trained_with$soilraster))){
        notintrainlevels <- levels(predstackdf$soilraster)[which(!levels(predstackdf$soilraster) %in% levels(trained_with$soilraster))]
        remove <- predstackdf$soilraster[predstackdf$soilraster %in% notintrainlevels]
        predstackdf$soilraster[predstackdf$soilraster %in% notintrainlevels] <- NA
        predstackdf$soilraster <- droplevels(predstackdf$soilraster)
        print(levels(predstackdf$soilraster))
        print("because these soil factor levels were not within the training boundaries, the following samples will be set to NA")
        print(table(droplevels(remove)))
      }
      
      compcaseloc <- which(complete.cases(predstackdf))
      predstackdf_comp <- predstackdf[complete.cases(predstackdf),]
      

      # to equalize classes of training and test set as suggested in  https://stackoverflow.com/questions/24829674/r-random-forest-error-type-of-predictors-in-new-data-do-not-match 
      predstackdf_comp <- rbind(trained_with[1, ] , predstackdf_comp)
      predstackdf_comp <- predstackdf_comp[-1,]

      pred <- predict(model_final, predstackdf_comp)
      
      allpix <- predstack$TeAqNum[]
      allpix <- NA
      allpix[compcaseloc] <- pred
      
      pred <- predstack$dem
      names(pred) <- "downscaled"
      pred[] <- allpix
      
      if(writeRastertoOutdir==TRUE){
              raster::writeRaster(pred, paste0(outdir, "dsc_", modname, add, outformat),
                          overwrite=T)
      }
      return(pred)
}


```


```{r examples-downscale}
library(raster)
library(lubridate)
library(solrad)
library(downscaleLST.MDV)


exdate <- as.POSIXct("2018-11-12 13:50:00", tz="UTC")
ex2 <- raster::stack("inst/ex_2_all_layers.grd")

##########  load the model ###############################################
load(system.file("final_model_downscaling_LST_MDV.RData", package = "downscaleLST.MDV"))

##########  make ia and put into one stack  ###############################
ia_ex2 <- make_ia(M_date = exdate, outdir = here::here("ex_output"),
                     sl=ex2$slope, as=ex2$aspect, tmplt=ex2$template)
names(ia_ex2) <- "ia"
ex2_ia <- raster::stack(ex2, ia_ex2)


dsc_ex2 <- downscaleFun(MOD=ex2_ia$Mbilin,
                        aux=ex2_ia[[c( "dem","slope","aspect","soilraster","landcoverres")]],
                        ia=ex2_ia$ia, 
                        template=ex2_ia$template, 
                        modname="MYD11_L2.A2018316.1350.006",
                        outdir = "ex_output/",
                        writeRastertoOutdir = TRUE,
                        add="_ex2",
                        outformat=".grd")

comp_ex2 <- raster::stack(ex2$Mbilin, ex2$Mngb, dsc_ex2, ex2$L)
raster::writeRaster(comp_ex2, "inst/comp_ex2.grd",overwrite=T)

```

# make_dsc_figure: Make a figure for comparing MODIS to downscaled and reference scene

```{r function-figure}
#' make a figure for comparing MODIS to downscaled and reference scene
#'
#' this function was adapted from ggAOA by Ludwigm6 https://github.com/Ludwigm6/mlutils/blob/b8fab319f8054df6bb4595bd8241a2d50c7753e3/R/plot_aoa.R
#' @param LST raster stack containing either MODIS and downscaled MODIS scene or the same plus a reference scene, preferably in this order, will be named by layer names
#' @param ex list containing extents that shall be plotted into the overview of the scene and displayed in a zoomed form below 
#' @param c character string with R colours names to use for smaller extent headings and boxes in main figure, should be same length as list in ex
#' @return
#' 
#' @export
#'
#' @examples

make_dsc_figure = function(LST, aoa=NULL, LST_name = "LST", ex=NULL, c=NULL,
                 limlow=min(LST[],na.rm=T), limhigh=max(LST[],na.rm=T),
                 legendposition="right",exsize=1, limitx=NULL, limity=NULL){
  
  if(!is.null(aoa)){
    # mask aoa
    aoa[aoa == 1] = NA
    
    base <- RStoolbox::ggR(LST, geom_raster = TRUE)+
      ggplot2::scale_fill_gradientn(name = LST_name, colors = viridis::inferno(50),
                                    na.value="transparent",
                                    limits=c(limlow,limhigh))+
      ggnewscale::new_scale_fill()+
      RStoolbox::ggR(aoa, ggLayer = TRUE, forceCat = TRUE, 
                     geom_raster = TRUE, alpha = NA)+
      ggplot2::scale_fill_manual(name = "Out of AOA", 
                                 values = c("0" = "grey70"), 
                                 na.value = "transparent")+
      ggplot2::coord_equal(expand = FALSE,xlim = limitx, ylim=limity)+
      ggplot2::scale_y_continuous(n.breaks=3)+
      ggplot2::scale_x_continuous(n.breaks=3)+
      theme_map()+ggplot2::theme(legend.position=legendposition)
    
  } else {
    base <- RStoolbox::ggR(LST, geom_raster = TRUE)+
      ggplot2::scale_fill_gradientn(name = LST_name, colors = viridis::inferno(50),
                                    na.value="transparent",
                                    limits=c(limlow,limhigh))+
      ggplot2::coord_equal(expand = FALSE, xlim = limitx, ylim=limity)+
      ggplot2::scale_y_continuous(n.breaks=3)+
      ggplot2::scale_x_continuous(n.breaks=3)+
      theme_map()+ggplot2::theme(legend.position=legendposition)
  }
  

  
    if(!is.null(ex)){ # if extents are supplied, make subplot and extents 
  
    
    explotlist <- lapply(seq(ex), function(i){
      ggplot2::geom_polygon(data=data.frame(x=c(raster::extent(ex[[i]])[c(1,1)], raster::extent(ex[[i]])[c(2,2)]),
                                   y=c(raster::extent(ex[[i]])[3:4], raster::extent(ex[[i]])[c(4,3)])),
                   aes(x = x, y = y), colour = c[i], fill=NA, size = exsize)
    })
    
    base <- base+explotlist
    
    } 
  
  return(base)
}


theme_map = function(y90 = TRUE, ts=10){
  t = ggplot2::theme_void()+
    ggplot2::theme(axis.line =  ggplot2::element_blank(),
    axis.title = ggplot2::element_blank(),
    axis.text = ggplot2::element_text(color = "black", size = ts),
    axis.ticks =  ggplot2::element_blank(),
    panel.grid = ggplot2::element_blank(),
    panel.border =  ggplot2::element_blank(),
    panel.background = ggplot2::element_rect(fill = "white"),
    legend.position = "right")
  
  if(y90){
    t = t + ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 90, hjust = 0.5))
  }
  
  
  return(t)
}

```



```{r examples-figure}
library(rasterVis)
library(sf)
library(viridis)
library(gridExtra)


comp_ex2 <- raster::stack("inst/comp_ex2.grd") # from example for downscaleFun

aoa <- raster::raster(list.files("inst/", pattern="AOA", full.names = T))

# this is the result of applying the 
aoa_ex2 <- raster::crop(aoa, comp_ex2)
aoa_ex2[aoa_ex2 == 1] = NA # set pixels that are within AOA to NA so  they don't overlap valid LST data

ex2 <- raster::stack(comp_ex2[[2:4]], aoa_ex2)
names(ex2) <- c("MODIS_LST", "downscaled_LST","Landsat_LST", "AOA")

# # stack original LST, downscaled LST and reference LST 
# M_dsc_L_stack <- raster::stack(full$Mngb, dsc_M, full$L, aoa)

# ex2 <- crop(M_dsc_L_stack, extent(dsc_ex2))

############### PLOTTING ###################################################
# 
# ### full scene ################
# dscplot <- make_dsc_figure(LST = M_dsc_L_stack$downscaled_LST,  LST_name = "LST",
#                  aoa=M_dsc_L_stack$AOA, ex=list(ex2, ex4), c=c("navy","chartreuse4"), 
#                  limlow = -40, limhigh = 0, exsize = 1.5, 
#                  limitx = c(350500, 500000), limity = c(-1400000, -1200000),legendposition = "none"
#                  )
# mplot <- make_dsc_figure(LST = M_dsc_L_stack$MODIS_LST,  LST_name = "MODIS LST",
#                  ex=list(ex2, ex4), c=c("navy","chartreuse4"), 
#                  limlow = -40, limhigh = 0, exsize = 1.5, 
#                  limitx = c(350500, 500050), limity = c(-1400050, -1200000),legendposition = "none"
#                )
# lplot <- make_dsc_figure(LST = M_dsc_L_stack$Landsat_LST,  LST_name = "Landsat LST",
#                ex=list(ex2, ex4), c=c("navy","chartreuse4"), 
#                limlow = -40, limhigh = 0, exsize = 1.5, 
#                limitx = c(350500, 500050), limity = c(-1400050, -1200000),legendposition = "none"
# )
# 
# plotlist <- list(mplot, dscplot, lplot)
# 
# margin = theme(plot.margin = unit(c(0.7,0.7,0.4,0.4), "cm"))
# 
# # png(paste0(figurepath, "full_example_scene_with_aoa.png"),
# #     units="cm", width=21, height=9, res=300)
# grid.arrange(
#   grobs = lapply(plotlist, "+", margin),
#   layout_matrix = rbind(c(1,2,3)))
# # dev.off()



### example 2 #################
library(ggplot2)
margin = ggplot2::theme(plot.margin = unit(c(0.4,0.4,0.4,0.4), "cm"))

dscplot <- make_dsc_figure(LST = ex2$downscaled_LST,  LST_name = "LST",
                 aoa=ex2$AOA, ex=list(ex2), c=c("navy"), 
                 limlow = -26, limhigh = -3, legendposition = "none",
                 exsize = 4)

mplot <- make_dsc_figure(LST = ex2$MODIS_LST,  LST_name = "MODIS LST",
               limlow = -26, limhigh = -3,ex=list(ex2), c=c("navy"), legendposition = "none",
               exsize = 4)

lplot <- make_dsc_figure(LST = ex2$Landsat_LST,  LST_name = "Landsat LST",
               limlow = -26, limhigh = -3,ex=list(ex2), c=c("navy"), legendposition = "none",
               exsize = 4)

plotlist <- list(mplot, dscplot, lplot)


gridExtra::grid.arrange(
  grobs = lapply(plotlist, "+", margin),
  layout_matrix = rbind(c(1,2,3)))
dev.off()

# with legend
dscplot_with_legend <- make_dsc_figure(LST = ex2$downscaled_LST,  LST_name = "LST",
                 aoa=ex2$AOA, ex=list(ex2), c=c("navy"), 
                 limlow = -26, limhigh = -3,
                 exsize = 2)

dscplot_with_legend
```


## To do: 

* Make function that gets recent scenes or specific scenes from MOD/MYD and warps all alone.
* implement tests 


# There can be development actions

These will be included in the `dev_history.R` file of your package, but won't be direct part of it.

- open project
- remove vignettes
- run

```{r development-1, eval=FALSE}
# Run but keep eval=FALSE to avoid infinite loop
# Execute in the console directly

vf <- list.files(here::here("vignettes"), full.names=T)
file.remove(vf)
fusen::inflate(rmd = "dev/dev_history.Rmd",name = "How_to_use_downscaleLST.MDV", check = TRUE)
```


# Inflate your package

You're one inflate from paper to box.
Build your package from this very Rmd using `fusen::inflate()`

- Verify your `"DESCRIPTION"` file has been updated
- Verify your function is in `"R/"` directory
- Verify your test is in `"tests/testthat/"` directory
- Verify this Rmd appears in `"vignettes/"` directory

