---
title: "How_to_use_downscaleLST.MDV"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How_to_use_downscaleLST.MDV}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(downscaleLST.MDV)
```

```{r setup_wd, include = FALSE}
knitr::opts_knit$set(root.dir = '../')
```

Package was generated with the help of package fusen by Sébastien Rochette. 


# Read data

The downscaling model contained in this package is called "final_model_downscaling_LST_MDV.RData". 

Auxiliary raster data are in aux.tif and an exemplary MODIS scene to downscale is in MOD11_L2_example.tif, it's from XXX (ID=XXX).

There is also a template file (template.tif) for the 30m resolution and slope.tif and aspect.tif (in degrees) to calculate the incidence angle.

<!-- Store your dataset in a directory named "inst/" at the root of your project -->
<!-- Use it for your tests in this Rmd thanks to `load_all()` to make it available
and `system.file()` to read it in your examples 
-->


# make_ia: Calculate the incidence angle for the MODIS scene

The example is based on a small extent, for which template, slope and aspect are provided in the package as "ex_temp.grd", "ex_slope.grd" and "ex_aspect.grd". They are prepared as follows: 


For the full area of interest grid, call: 

full <- raster(system.file("full_size_grids_all_layers.grd", package = "downscaleLST.MDV"))
ex <- raster(system.file("ex_2_all_layers.grd", package = "downscaleLST.MDV"))



```{r examples}

exdate <- as.POSIXct("2018-11-12 13:50:00", tz="UTC")
full <- raster::stack("inst/full_size_grids_all_layers.grd")
ex2 <- raster::stack("inst/ex_2_all_layers.grd")

# for the example extent
ia <- make_ia(M_date = exdate, outdir = here::here("ex_day"), 
              sl=ex2$slope, as=ex2$aspect, tmplt=ex2$template)
# # for the full scene
# ia <- make_ia(M_date = exdate, outdir = here::here("ex_day"), 
#               sl=full$slope, as=full$aspect, tmplt=full$template)


```

# downscaleFun: Downscale MODIS scene 

```{r examples-downscale}
library(raster)
library(lubridate)
library(solrad)
library(downscaleLST.MDV)


exdate <- as.POSIXct("2018-11-12 13:50:00", tz="UTC")
ex2 <- raster::stack("inst/ex_2_all_layers.grd")

##########  load the model ###############################################
load(system.file("final_model_downscaling_LST_MDV.RData", package = "downscaleLST.MDV"))

##########  make ia and put into one stack  ###############################
ia_ex2 <- make_ia(M_date = exdate, outdir = here::here("ex_output"),
                     sl=ex2$slope, as=ex2$aspect, tmplt=ex2$template)
names(ia_ex2) <- "ia"
ex2_ia <- raster::stack(ex2, ia_ex2)


dsc_ex2 <- downscaleFun(MOD=ex2_ia$Mbilin,
                        aux=ex2_ia[[c( "dem","slope","aspect","soilraster","landcoverres")]],
                        ia=ex2_ia$ia, 
                        template=ex2_ia$template, 
                        modname="MYD11_L2.A2018316.1350.006",
                        outdir = "ex_output/",
                        writeRastertoOutdir = TRUE,
                        add="_ex2",
                        outformat=".grd")

comp_ex2 <- raster::stack(ex2$Mbilin, ex2$Mngb, dsc_ex2, ex2$L)
raster::writeRaster(comp_ex2, "inst/comp_ex2.grd",overwrite=T)

```

# make_dsc_figure: Make a figure for comparing MODIS to downscaled and reference scene

```{r examples-figure}
library(rasterVis)
library(sf)
library(viridis)
library(gridExtra)


comp_ex2 <- raster::stack("inst/comp_ex2.grd") # from example for downscaleFun

aoa <- raster::raster(list.files("inst/", pattern="AOA", full.names = T))

# this is the result of applying the 
aoa_ex2 <- raster::crop(aoa, comp_ex2)
aoa_ex2[aoa_ex2 == 1] = NA # set pixels that are within AOA to NA so  they don't overlap valid LST data

ex2 <- raster::stack(comp_ex2[[2:4]], aoa_ex2)
names(ex2) <- c("MODIS_LST", "downscaled_LST","Landsat_LST", "AOA")

# # stack original LST, downscaled LST and reference LST 
# M_dsc_L_stack <- raster::stack(full$Mngb, dsc_M, full$L, aoa)

# ex2 <- crop(M_dsc_L_stack, extent(dsc_ex2))

############### PLOTTING ###################################################
# 
# ### full scene ################
# dscplot <- make_dsc_figure(LST = M_dsc_L_stack$downscaled_LST,  LST_name = "LST",
#                  aoa=M_dsc_L_stack$AOA, ex=list(ex2, ex4), c=c("navy","chartreuse4"), 
#                  limlow = -40, limhigh = 0, exsize = 1.5, 
#                  limitx = c(350500, 500000), limity = c(-1400000, -1200000),legendposition = "none"
#                  )
# mplot <- make_dsc_figure(LST = M_dsc_L_stack$MODIS_LST,  LST_name = "MODIS LST",
#                  ex=list(ex2, ex4), c=c("navy","chartreuse4"), 
#                  limlow = -40, limhigh = 0, exsize = 1.5, 
#                  limitx = c(350500, 500050), limity = c(-1400050, -1200000),legendposition = "none"
#                )
# lplot <- make_dsc_figure(LST = M_dsc_L_stack$Landsat_LST,  LST_name = "Landsat LST",
#                ex=list(ex2, ex4), c=c("navy","chartreuse4"), 
#                limlow = -40, limhigh = 0, exsize = 1.5, 
#                limitx = c(350500, 500050), limity = c(-1400050, -1200000),legendposition = "none"
# )
# 
# plotlist <- list(mplot, dscplot, lplot)
# 
# margin = theme(plot.margin = unit(c(0.7,0.7,0.4,0.4), "cm"))
# 
# # png(paste0(figurepath, "full_example_scene_with_aoa.png"),
# #     units="cm", width=21, height=9, res=300)
# grid.arrange(
#   grobs = lapply(plotlist, "+", margin),
#   layout_matrix = rbind(c(1,2,3)))
# # dev.off()



### example 2 #################
library(ggplot2)
margin = ggplot2::theme(plot.margin = unit(c(0.4,0.4,0.4,0.4), "cm"))

dscplot <- make_dsc_figure(LST = ex2$downscaled_LST,  LST_name = "LST",
                 aoa=ex2$AOA, ex=list(ex2), c=c("navy"), 
                 limlow = -26, limhigh = -3, legendposition = "none",
                 exsize = 4)

mplot <- make_dsc_figure(LST = ex2$MODIS_LST,  LST_name = "MODIS LST",
               limlow = -26, limhigh = -3,ex=list(ex2), c=c("navy"), legendposition = "none",
               exsize = 4)

lplot <- make_dsc_figure(LST = ex2$Landsat_LST,  LST_name = "Landsat LST",
               limlow = -26, limhigh = -3,ex=list(ex2), c=c("navy"), legendposition = "none",
               exsize = 4)

plotlist <- list(mplot, dscplot, lplot)


gridExtra::grid.arrange(
  grobs = lapply(plotlist, "+", margin),
  layout_matrix = rbind(c(1,2,3)))
dev.off()

# with legend
dscplot_with_legend <- make_dsc_figure(LST = ex2$downscaled_LST,  LST_name = "LST",
                 aoa=ex2$AOA, ex=list(ex2), c=c("navy"), 
                 limlow = -26, limhigh = -3,
                 exsize = 2)

dscplot_with_legend
```

## To do: 

* Make function that gets recent scenes or specific scenes from MOD/MYD and warps all alone.
* implement tests 



# There can be development actions

These will be included in the `dev_history.R` file of your package, but won't be direct part of it.

- open project
- remove vignettes
- run


# Inflate your package

You're one inflate from paper to box.
Build your package from this very Rmd using `fusen::inflate()`

- Verify your `"DESCRIPTION"` file has been updated
- Verify your function is in `"R/"` directory
- Verify your test is in `"tests/testthat/"` directory
- Verify this Rmd appears in `"vignettes/"` directory


